---
title: "WC EDA"
author: "Thomas Sato"
date: "2024-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages
```{r}
library(tidyverse)
library(sf)
library(readxl)
```

# Read in data
```{r}
# These are the shapefiles that we use for mapping the decision units.
wc_data <- st_read("320001056-18 (App. C Grids)5.dxf")
st_crs(wc_data) <- 2261
```
```{r}
# This is the data from the study, includes all concentrations of metals and other chemicals.
metals <- read_excel("WillametteCoveData.xlsx", sheet = "Metals by EPA Method 6020B")
dioxins_furans <- read_excel("WillametteCoveData.xlsx", sheet = "Dioxins Furans by EPA Method 16")
PAHs <- read_excel("WillametteCoveData.xlsx", sheet = "PAHs and Dibenzofuran by EPA Me")
PCBs <- read_excel("WillametteCoveData.xlsx", sheet = "PCBs by EPA Method 8082A")

# Hot spot levels
hot_spot <- read_excel("WillametteCoveData.xlsx", sheet = "Hot Spot Levels")
```

# Data Cleaning

## For map data:
```{r}
# Filters out one 3D LINESTRING that broke the graphic.
# The 00 Grid Areas layer includes boundaries of decision units.
decision_units <- filter(wc_data, Layer == "00 Grid Areas" & EntityHandle != "17ADF")
```

Necessary for identifying the decision units and labeling as well.
```{r}
decision_units <- decision_units %>%
  mutate(Text = recode(EntityHandle,
                       "B66B" = "DU-1",
                       "B654" = "DU-2",
                       "B66C" = "DU-3",
                       "B653" = "DU-4",
                       "B66A" = "DU-5",
                       "B655" = "DU-6",
                       "B669" = "DU-7",
                       "B668" = "DU-8",
                       "B667" = "DU-9",
                       "B664" = "DU-10",
                       "B662" = "DU-11",
                       "B65E" = "DU-12",
                       "B663" = "DU-13",
                       "B66E" = "DU-14",
                       "B65F" = "DU-15",
                       "17B12" = "DU-16",
                       "B657" = "DU-17",
                       "B674" = "DU-18",
                       "B66D" = "DU-19",
                       "14FF5" = "DU-20",
                       "B673" = "DU-21",
                       "14FFD" = "DU-22",
                       "15005" = "DU-23",
                       "B661" = "DU-24",
                       "B666" = "DU-25",
                       "B665" = "DU-26",
                       "B660" = "DU-27",
                       "D1E8" = "DU-28",
                       "D1F1" = "DU-29",
                       "D1F0" = "DU-30",
                       "17E12" = "DU-31",
                       "17E18" = "DU-31",
                       "D1F7" = "DU-32",
                       "B676" = "DU-33",
                       "B65A" = "DU-34",
                       "B65B" = "DU-35",
                       "1802C" = "DU-36",
                       "B66F" = "DU-36",
                       "B670" = "DU-37",
                       "B658" = "DU-38",
                       "B60F" = "DU-38",
                       "BE09" = "DU-39",
                       "EA3A" = "DU-39",
                       "EA3B" = "DU-40",
                       "D1EA" = "DU-41",
                       "17E0A" = "DU-43",
                       "18029" = "DU-44",
                       "BE89" = NA_character_))
```

Need to convert LINESTRINGs to POLYGONs in order to be able to fill.
```{r}
linestrings <- decision_units$geometry

polygons <- lapply(linestrings, function(ls) {
  
  coords <- st_coordinates(ls)
  
  # Close the LINESTRING
  closed_coords <- rbind(coords, coords[1, ])
  
  # Create POLYGON
  st_polygon(list(closed_coords))
  
})

# Set Coordinate Reference System to be able to plot correctly
polygons_sf <- st_sfc(polygons, crs = 2261)

concentrations_polygons <- decision_units %>%
  mutate(geometry = polygons_sf)

# Removes unnecessary Z coordinate of geometries
concentrations_polygons$geometry <- st_zm(concentrations_polygons$geometry)
```


## For soil sample data:
```{r}
# Constants of columns for cleaning data.

dioxin_furan_cols <- c(
  "2,3,7,8-TCDD",
  "1,2,3,7,8-PeCDD",
  "1,2,3,4,7,8-HxCDD",
  "1,2,3,6,7,8-HxCDD",
  "1,2,3,7,8,9-HxCDD",
  "1,2,3,4,6,7,8-HpCDD",
  "OCDD",
  "2,3,7,8-TCDF",
  "1,2,3,7,8-PeCDF",
  "2,3,4,7,8-PeCDF",
  "1,2,3,4,7,8-HxCDF",                     
  "1,2,3,6,7,8-HxCDF",                     
  "2,3,4,6,7,8-HxCDF",                     
  "1,2,3,7,8,9-HxCDF",                      
  "1,2,3,4,7,8-HpCDF",                       
  "1,2,3,4,7,8,9-HpCDF",                    
  "OCDF",                                             
  "Total TCDD",                                       
  "Total PeCDD",                                  
  "Total HxCDD",                               
  "Total HpCDD",                                  
  "Total TCDF",                          
  "Total PeCDF",                                   
  "Total HxCDF",                                    
  "Total HpCDF",                                     
  "Total D/F TEQ (Mammal)")

PAHs_cols <- c(
  "Acenaphthene",
  "Acenaphthylene",
  "Anthracene",
  "Benz(a) anthracene",
  "Benzo(a) pyrene",
  "Benzo(b) fluoranthene",
  "Benzo(k) fluoranthene",
  "Benzo(g,h,i) perylene",
  "Chrysene",
  "Dibenz(a,h) anthracene",              
  "Fluoranthene",                                                       
  "Fluorene",                                                                 
  "Indeno(1,2,3-cd)pyrene",                                 
  "1-Methylnaphthalene",                                          
  "2-Methylnaphthalene",                                     
  "Naphthalene",                                                       
  "Phenanthrene",                                                   
  "Pyrene",                                                               
  "Dibenzofuran",
  "Total HPAH",                                                          
  "Total LPAH",                                                   
  "cPAHs (BaP Eq.)")

PCBs_cols <- c(
  "Aroclor 1016",
  "Aroclor 1221",
  "Aroclor 1232",
  "Aroclor 1242",
  "Aroclor 1248",
  "Aroclor 1254",
  "Aroclor 1260",
  "Aroclor 1262",
  "Aroclor 1268",
  "Total PCB Aroclors")
```

Converting to long format.
```{r}
metals_long <- metals %>%
  select(-"Metals by EPA Method 6020B") %>%
  pivot_longer(
    cols = c(Antimony, Arsenic, Chromium, Copper, Lead, Mercury, Nickel, Selenium, Zinc),  
    names_to = "Chemical",                      
    values_to = "Concentration"                    
  ) %>%
  pivot_longer(
    cols = ends_with("_detection"),         
    names_to = "Chemical_detection",              
    values_to = "Detection"                       
  ) %>%
  mutate(
    Chemical_detection = str_replace(Chemical_detection, "_detection", "")
  ) %>%
  filter(Chemical == Chemical_detection) %>%
  filter(`Sample Type...6` != "Mean") %>% # Getting rid of aggregate calculations
  select(-Chemical_detection, -`Sample Type...6`) %>%
  rename("Sample Type" = "Sample Type...3")
```

```{r}
dioxins_furans_long <- dioxins_furans %>%
  select(-"Dioxins/Furans by EPA Method 1613B") %>%
  pivot_longer(
    cols = all_of(dioxin_furan_cols),
    names_to = "Chemical",                      
    values_to = "Concentration"                    
  ) %>%
  pivot_longer(
    cols = ends_with("_detection"),         
    names_to = "Chemical_detection",              
    values_to = "Detection"                       
  ) %>%
  mutate(
    Chemical_detection = str_replace(Chemical_detection, "_detection", "")
  ) %>%
  filter(Chemical == Chemical_detection) %>%      
  select(-Chemical_detection)
```

```{r}
PAHs_long <- PAHs %>%
  select(-"PAHs and Dibenzofuran by EPA Method 8270E-SIM") %>%
  pivot_longer(
    cols = all_of(PAHs_cols),
    names_to = "Chemical",                      
    values_to = "Concentration"                    
  ) %>%
  pivot_longer(
    cols = ends_with("_detection"),         
    names_to = "Chemical_detection",              
    values_to = "Detection"                       
  ) %>%
  mutate(
    Chemical_detection = str_replace(Chemical_detection, "_detection", "")
  ) %>%
  filter(Chemical == Chemical_detection) %>%      
  select(-Chemical_detection)  
```

```{r}
PCBs_long <- PCBs %>%
  select(-"PCBs by EPA Method 8082A") %>%
  pivot_longer(
    cols = all_of(PCBs_cols),
    names_to = "Chemical",                      
    values_to = "Concentration"                    
  ) %>%
  pivot_longer(
    cols = ends_with("_detection"),         
    names_to = "Chemical_detection",              
    values_to = "Detection"                       
  ) %>%
  mutate(
    Chemical_detection = str_replace(Chemical_detection, "_detection", "")
  ) %>%
  filter(Chemical == Chemical_detection) %>%      
  select(-Chemical_detection)
```

Union of all long datasets together for one big one
```{r}
concentrations <- rbind(metals_long, dioxins_furans_long, PAHs_long, PCBs_long) %>%
  filter(`Sample Depth (feet bgs)` %in% c("0-1", "1-2", "2-3")) %>% # Removing Field Replicates and keeping original data.
  filter(!str_detect(`Sample ID`, "M$")) %>% # Filtering out mean samples (ends with M)
  mutate(`Decision Unit` = substr(`Decision Unit`, 1, 5)) # Need first 5 characters (i.e. DU-44) 
```

Joining together decision units and concentrations
```{r}
concentrations_geometries <- left_join(concentrations_polygons,
                                       concentrations,
                                       by = c("Text" = "Decision Unit"))
```

Hot spot levels
```{r}
# Recoding values for later join
hot_spot$Page23RDIWorkPlan <- recode(hot_spot$Page23RDIWorkPlan,
                                     "Dioxin/Furan TEQ" = "Total D/F TEQ (Mammal)",
                                     "Total PCBs" = "Total PCB Aroclors")
```

Testing correction for hot spots
```{r}
hot_spot_accurate <- hot_spot %>%
  mutate(
    `ISM PRG` = case_when(
      Page23RDIWorkPlan == "Total HPAH" ~ `ISM PRG` * 1000,
      Page23RDIWorkPlan == "Total LPAH" ~ `ISM PRG` * 1000,
      Page23RDIWorkPlan == "Dibenzofuran" ~ `ISM PRG` * 1000,
      Page23RDIWorkPlan == "Total PCB Aroclors" ~ `ISM PRG` * 1000,
      Page23RDIWorkPlan == "Total D/F TEQ (Mammal)" ~ `ISM PRG` * 1000000,
      TRUE ~ `ISM PRG`
    ),
    `ISM Hot Spot Level` = case_when(
      Page23RDIWorkPlan == "Total HPAH" ~ `ISM Hot Spot Level` * 1000,
      Page23RDIWorkPlan == "Total LPAH" ~ `ISM Hot Spot Level` * 1000,
      Page23RDIWorkPlan == "Dibenzofuran" ~ `ISM Hot Spot Level` * 1000,
      Page23RDIWorkPlan == "Total PCB Aroclors" ~ `ISM Hot Spot Level` * 1000,
      Page23RDIWorkPlan == "Total D/F TEQ (Mammal)" ~ `ISM Hot Spot Level` * 1000000,
      TRUE ~ `ISM Hot Spot Level`
    )
  )
```



Basic plot of area of interest.
```{r}
ggplot(data = decision_units) +
  geom_sf() +
  theme_minimal() +
  labs(title = "Willamette Cove")
```

### Antimony concentrations for the Willamette Cove by depth

```{r}
antimony <- concentrations_geometries %>%
  filter(Chemical == "Antimony") %>%
  mutate(concentration_binned = cut(Concentration, 
                                    breaks = c(0, 0.5, 0.75, 1, Inf),
                                    labels = c("< 0.50", "0.50 - 0.75", "0.75 - 1.00", "> 1.00")))
```

```{r}
heavy_metals_plot <- ggplot(antimony) +
  geom_sf(aes(fill = concentration_binned)) +
  scale_fill_manual(values = c("#ADD8E6", "#FA8072", "#DE6055", "#C24039")) +
  theme_minimal() +
  facet_wrap(~ `Sample Depth (feet bgs)`,
             ncol = 1,
             strip.position = "left",
             labeller = as_labeller(function(x) paste(x, "\nft bgs"))) +
  labs(title = "Antimony Makes Its Mark in the East Parcel",
       subtitle = "Higher concentrations of antimony found in the northeast corner of the East Parcel, \nalong with parts of the Central Parcel",
       fill = "Concentration of \nAntimony (mg/kg)",
       caption = "Source: Willamette Cove RDI Evaluation Report")
```

```{r}
(
heavy_metals_plot <- heavy_metals_plot +
  theme(
    panel.spacing = unit(-3, "lines"),
    panel.grid = element_blank(),
    strip.text = element_text(margin = margin(t = 10, b = 75)),
    strip.text.y.left = element_text(size = 16, angle = 0),
    plot.title = element_text(size = 18),
    axis.text = element_blank()
  )
)
```

Ideally, this graph would be made for each contaminant of concern, and then we would be able to change between them easily with the interactive Shiny App.

### Contaminant of Concern concentrations for Decision Unit 1:

```{r}
cocs <- c("Antimony", "Arsenic", "Chromium", "Copper", "Lead", "Mercury", "Nickel", "Selenium", "Zinc", "Total HPAH", "Total LPAH", "Dibenzofuran", "Total PCB Aroclors", "Total D/F TEQ (Mammal)")

DU_1 <- concentrations %>%
  filter(`Decision Unit` == "DU-13" & Chemical %in% cocs)
```

```{r}
DU_1_HS <- left_join(DU_1, hot_spot_accurate, by = c("Chemical" = "Page23RDIWorkPlan"))
DU_1_HS <- mutate(DU_1_HS, exceedance = case_when(
  Concentration > `ISM Hot Spot Level` ~ "Exceeds Hot Spot Level",
  Concentration > `ISM PRG` ~ "Exceeds PRG",
  TRUE ~ "Does not Exceed"
))
```

```{r}
# Figure out a way to order the chemicals for this graphic
ordering <- DU_1_HS %>%
  group_by(Chemical, exceedance) %>%
  summarize(weight = n()) %>%
  mutate(weight = case_when(
    exceedance == "Exceeds Hot Spot Level" ~ weight * 3,
    exceedance == "Exceeds PRG" ~ weight * 2,
    TRUE ~ weight
  )) %>%
  group_by(Chemical) %>%
  summarize(weight = sum(weight))
```

```{r}
DU_1_HS <- left_join(DU_1_HS, ordering)
DU_1_HS$exceedance <- factor(DU_1_HS$exceedance, levels = c("Exceeds Hot Spot Level", "Exceeds PRG", "Does not Exceed"))
```


```{r}
DU_1_plot <- ggplot(DU_1_HS) +
  geom_tile(aes(`Sample Depth (feet bgs)`, reorder(Chemical, weight), fill = exceedance)) +
  theme_minimal() +
  labs(title = "Which chemicals exeed ecological \nhot spot values in DU-1?",
       subtitle = "PRG = Preliminary Remediation Goal",
       y = "Contaminant of Concern",
       fill = "",
       caption = "Source: Willamette Cove RDI Evaluation Report") +
  scale_fill_manual(values = c("Does not Exceed" = "#deebf7",
                               "Exceeds PRG" = "#9ecae1",
                               "Exceeds Hot Spot Level" = "#3182bd"))
```

```{r}
(
DU_1_plot <- DU_1_plot +
  theme(
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 12)
  )
)
```



