---
title: "WC EDA Basic Plots"
author: "Kevin Kiser"
date: "2024-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "WC EDA Plots"
author: "Kevin Kiser"
date: "2024-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```



# Data Import

```{r}

results_summary <- read_excel("WillametteCoveData.xlsx", sheet = "Results Summary Ecological")
avg_concentrations <- read_excel("WillametteCoveData.xlsx", sheet = "Average Concentrations By Depth")
exceedance_ratios <- read_excel("WillametteCoveData.xlsx", sheet = "Exceedance Ratios")


# concentrations from cleaned 
# "Metals by EPA Method 6020B"
# "Dioxins Furans by EPA Method 16"
# "PAHs and Dibenzofuran by EPA Me"
# "PCBs by EPA Method 8082A"

concentrations <- read.csv("concentrations.csv")
```

# Create Heatmap plot function
creates a heatmap for any given; chemical, location, and Threshold (ecological/human)

```{r}

create_heatmap <- function(chemical, location, affix) {
  
  red_palette <- colorRampPalette(c("salmon", "darkred"))(5)

  threshold_string <- if_else(affix == "_Eco", "Ecological", "Human Health")

  # Process the exceedance ratios and create the heatmap
  heatmap_plot <- exceedance_ratios %>%
    rename(`Sample Depth` = `Sample Depth (feet bgs)`) %>%
    filter(`Sample Location` == location) %>%
    select(`Decision Unit`, `Sample Depth`, matches(paste0("^", chemical, ".*", affix, "$"))) %>%
    pivot_wider(
      names_from = `Sample Depth`,  
      values_from = matches(paste0("^", chemical, affix, "$"))
    ) %>%
    pivot_longer(-`Decision Unit`, names_to = "Sample Depth", values_to = "Exceedance_Ratio") %>%
    mutate(`Sample Depth` = factor(`Sample Depth`, levels = sort(unique(`Sample Depth`), decreasing = TRUE)),
           # Set color_range as a factor with specific levels for ordered legend
           color_range = factor(case_when(
             Exceedance_Ratio < 1 ~ "(< 1.0) Acceptable",
             Exceedance_Ratio < 2 ~ "(1.0 - 2.0)",
             Exceedance_Ratio < 3 ~ "(2.0 - 3.0)",
             Exceedance_Ratio < 4 ~ "(3.0 - 4.0)",
             Exceedance_Ratio >= 4 ~ "(> 4.0)"
           ), levels = c("(< 1.0) Acceptable", "(1.0 - 2.0)", "(2.0 - 3.0)", "(3.0 - 4.0)", "(> 4.0)"))) %>%
    
    ggplot(aes(x = `Decision Unit`, y = `Sample Depth`, 
               fill = color_range)) +  # Fill based on the new Range column
    
    geom_tile(color = "white", linewidth = 2) +
    
    geom_text(aes(label = round(Exceedance_Ratio, 1)), color = "black", size = 4) +
    
    labs(title = paste("Contaminant: ", chemical, "\n", threshold_string, " Threshold", "\n", location, sep=""),
         x = "Decision Unit",
         y = "Sample Depth (feet)",
         fill = "Concentration \nper\nThreshold") +  # Legend title
         
    theme_minimal() +
    
    scale_fill_manual(values = c(
      "(< 1.0) Acceptable" = "lightblue",
      "(1.0 - 2.0)" = red_palette[1],
      "(2.0 - 3.0)" = red_palette[2],
      "(3.0 - 4.0)" = red_palette[3],
      "(> 4.0)" = red_palette[4]
    )) +
    coord_fixed() +
    
    # Tilt x-axis labels by 90 degrees
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Return the heatmap plot
  return(heatmap_plot)
}


```


# Heatmap Example Calls

```{r}
create_heatmap("Lead", "East Parcel", "_Eco")
create_heatmap("Antimony", "Central Parcel", "_HH")
create_heatmap("Lead", "West Parcel", "_Eco")

```

```{r}

chemical_list <- c("Antimony", "Arsenic", "Chromium", "Copper", "Lead", "Mercury", "Nickel", "Selenium", "Zinc", "Total TEQ", "Dibenzofuran", "Total HPAH", "Total LPAH", "Total PCBs", "cPAHs")

parcel_list <- c("West Parcel", "Central Parcel", "Central Parcel Concrete Slab", "East Parcel", "East Parcel Concrete Slab", "East Parcel Soil Beams")

```


# Bar Plot Concentration Function 
Creates a plot of the highest chemical concentrations for any given DU

```{r}

du_concen_plot <- function(du) {
  chemical_list <- c(
    "Antimony", "Arsenic", "Chromium", "Copper", "Lead", "Mercury", "Nickel", "Selenium", "Zinc", 
    "OCDD", "OCDF", "Total TCDD", "Total PeCDD", "Total HxCDD", "Total HpCDD", 
    "Total TCDF", "Total PeCDF", "Total HxCDF", "Total HpCDF", "Total D/F TEQ (Mammal)", 
    "Acenaphthene", "Acenaphthylene", "Anthracene", "Benz(a) anthracene", 
    "Benzo(a) pyrene", "Benzo(b) fluoranthene", "Benzo(k) fluoranthene", "Benzo(g,h,i) perylene", 
    "Chrysene", "Dibenz(a,h) anthracene", "Fluoranthene", "Fluorene", "Indeno(1,2,3-cd)pyrene", 
    "1-Methylnaphthalene", "2-Methylnaphthalene", "Naphthalene", "Phenanthrene", 
    "Pyrene", "Dibenzofuran", "Total HPAH", "Total LPAH", "cPAHs (BaP Eq.)", "Total PCB Aroclors"
  )
  
  # Define the color palette
  palette2 <- c(
    "0-1" = "purple",
    "1-2" = "lightblue",
    "2-3" = "blue"
  )
  
  # Create the plot and assign it to du_plot
  du_plot <- concentrations %>%
    filter(Chemical %in% chemical_list,
           Decision.Unit == du) %>%
    mutate(`Sample.Depth..feet.bgs.` = factor(case_when(
      str_detect(`Sample.Depth..feet.bgs.`, "0-1") ~ "0-1",
      str_detect(`Sample.Depth..feet.bgs.`, "1-2") ~ "1-2",
      str_detect(`Sample.Depth..feet.bgs.`, "2-3") ~ "2-3",
      TRUE ~ `Sample.Depth..feet.bgs.`)),
      Chemical = fct_reorder(Chemical, Concentration, .desc = FALSE)) %>%
    ggplot(aes(x = Chemical, y = Concentration, fill = `Sample.Depth..feet.bgs.`)) +
    geom_col() +
    labs(title = paste("Sampling Location:", du),
         subtitle = "Chemical Concentrations",
         x = "Chemical",
         y = "Concentration (mg/kg)",
         fill = "Sample Depth\n(feet)") +
    scale_fill_manual(values = palette2) +
    theme_minimal() +
    theme(axis.text.x = element_text(hjust = 1, size = 10),
          axis.text.y = element_text(size = 6)) +
    coord_flip()
  
  return(du_plot)
}

```

# Example Calls

```{r}
du_concen_plot("DU-3")
du_concen_plot("DU-41")
```

# Plot Function for average concentrations for each parcel

```{r}


avg_concen_parcel_plot <- function(parcel){
  palette2 <- c(
    "0-1" = "purple",
    "1-2" = "lightblue",
    "2-3" = "blue")

plot <- avg_concentrations %>%
  pivot_longer(
    cols = 3:18,               
    names_to = "chemical",    
    values_to = "concentration" 
  ) %>%
  pivot_wider(
    names_from = `Sample Depth (feet bgs)`,   
    values_from = concentration               
  ) %>%
  pivot_longer(
    cols = 3:5,
    names_to =  "Depth",
    values_to = "concentration"
  ) %>%
  filter(chemical != "Total PCB Aroclors_detection",
         `Sample Location` == parcel) %>%
  mutate(concentration = as.numeric(round(as.numeric(concentration), 2)),
         chemical = fct_reorder(chemical, concentration, .desc = FALSE)) %>%
  ggplot(aes(x=chemical, y=concentration, fill=Depth)) +
  geom_col() +
  labs(title = paste("Sampling Location:", parcel),
         subtitle = "Average Chemical Concentrations",
         x = "Chemical",
         y = "Concentration (mg/kg)",
         fill = "Sample Depth\n(feet)") +
  scale_fill_manual(values = palette2) +
  theme_minimal() +
  coord_flip()
return(plot)

}  

```

#Example Calls

```{r}
# all possible parcel inputs
# [1] "West Parcel"                 
# [2] "Central Parcel"              
# [3] "Central Parcel Concrete Slab"
# [4] "East Parcel"                 
# [5] "East Parcel Concrete Slabs"  
# [6] "East Parcel Soil Berms" 

avg_concen_parcel_plot("West Parcel")
avg_concen_parcel_plot("East Parcel Concrete Slabs")
```





